---
title: Coursera Practical Machine Learning Prediction Project
author: "by Vincent Kowalski"
output:
  html_document:
    fig_height: 9
    fig_width: 9
---

## Background  
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.  

In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.  

## Libraries

```{r results='hide', message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
```

## Download data
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  
## Create training and test data sets
After downloading the data from the data source, we can read the two csv files into two data frames.  
```{r, cache = T}
training_data <- read.csv("./data/pml-training.csv")
testing_data <- read.csv("./data/pml-testing.csv")

###Processing Data

```{r}
#Delete columns with NA 
training_data <- training_data[, colSums(is.na(testing_data)) == 0]
testing_data  <- testing_data[, colSums(is.na(testing_data)) == 0]

#Filter data
training_data  <- training_data[, -c(1:7)]
testing_data   <- testing_data[, -c(1:7)]

dim(training_data)
dim(testing_data)
```

## Create Training and Testing data sets

```{r}
#Divide training_data to sub_training_data and sub_testing_data (75% and 25% resp.)
part_ind           <- createDataPartition(y = training_data$classe, p = 0.75, list = FALSE,)
sub_training_data  <- training_data[part_ind,]
sub_testing_data   <- training_data[-part_ind,]
```

## Create Random Forest and Decision Tree models and do comparison

```{r}
#Random Forest Model
rand_forest_model <- randomForest(classe ~. , data = sub_training_data, method = "class")
pred_rand_forest  <- predict(rand_forest_model, sub_testing_data, type = "class")
res_rand_forest    <- confusionMatrix(pred_rand_forest, sub_testing_data$classe)

#Decision Tree Model
dec_tree_model  <- rpart(classe ~ ., data = sub_training_data, method = "class")
pred_dec_tree   <- predict(dec_tree_model, sub_testing_data, type = "class")
res_rand_tree   <- confusionMatrix(pred_dec_tree, sub_testing_data$classe)
 
prp(dec_tree_model)

res_comp <- data.frame(res_rand_forest$overall, res_rand_tree$overall)
res_comp
```

## Prediction
```{r}
#Predict outcome on the testing_data using random forest model
final_pred <- predict(rand_forest_model, testing_data, type = "class")
final_pred
```